---
title: "week2_exercise"
author: "Cyril Geistlich"
date: "2023-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r import libraries}
library("readr") 
library("sf") 
library("ggplot2") # to visualize data
library("gitcreds")
library("dplyr")
```
# Task 1: Import Data

```{r import data}

boar<- read_delim("data/wildschwein_BE_2056.csv", ",")
boar<- st_as_sf(boar, coords = c("E", "N"), crs = 2056, remove = FALSE)
```
# Task 2
```{r Number of Animals}
(unique(boar$TierName))
```
3 Animals were tracked

```{r delta time}
(delta_time <- as.integer(difftime(max(boar$DatetimeUTC),min(boar$DatetimeUTC),units = 'secs')))
```
```{r Timespan plot}
ggplot(boar, aes(DatetimeUTC, TierName)) +
  geom_point()
```
The cows were tracked during different time intervals, but with a lot of overlap. 

```{r timespan per cow}
boar <- boar |>
  group_by(TierName) |>
  mutate(diff_s = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC)))

#Time span for each cow
('Ruth')
ruth <- boar[boar$TierName == 'Ruth',]
(min(ruth$DatetimeUTC))
(max(ruth$DatetimeUTC))
('Sabi')
sabi <- boar[boar$TierName == 'Sabi',]
(min(sabi$DatetimeUTC))
(max(sabi$DatetimeUTC))
('Rosa')
rosa <- boar[boar$TierName == 'Rosa',]
(min(rosa$DatetimeUTC))
(max(rosa$DatetimeUTC))
```
The Animals were tracked from "2014-08-22 21:00:12 UTC" until "2015-07-27 11:00:14 UTC". 
The Animals were not tracked during the same periods of time. Ruth & Rosa were tracked for a shorter period of time. 

```{r sample rate}
ggplot(boar, aes(diff_s)) + 
  geom_histogram(binwidth = 10) + 
  lims(x = c(0,5000),
       y = c(0,5000))

```
The sample rate was not continouse. There seem to be three main sample rates. 

```{r sample rate per cow}
ruth |>
  filter(DatetimeUTC < "2014-12-9") |>
  ggplot(aes(DatetimeUTC, diff_s), colour = TierName) +
  geom_point() +
  geom_line()

sabi |>
  filter(DatetimeUTC < "2014-8-30") |>
  ggplot(aes(DatetimeUTC, diff_s), colour = TierName) +
  geom_point() +
  geom_line()

rosa |>
  filter(DatetimeUTC < "2014-11-20") |>
  ggplot(aes(DatetimeUTC, diff_s), colour = TierName) +
  geom_point() +
  geom_line()
```
The sample rate for each cow was different. 
Ruth had an sample rate around 900s approximately the first month. After that the sample rate was changed to about 3000-3500s/sample and during the nights the sample rate was further reduced to about 11000s/sample. 
Sabi shows a sample rate of about 900s/sample during the day and 11000s/sample during the night.
Rosa had a sample rate of about 900 for the first days and later the sample rate changes to the same day and night cycle as Sabi. 
There seem to be some outliers in the samplerate for each cow. 

```{r max min samplingrate per cow}

("Ruth")
(min(ruth$diff_s, na.rm = TRUE))
(max(ruth$diff_s, na.rm = TRUE))


("Sabi")
(min(sabi$diff_s, na.rm = TRUE))
(max(sabi$diff_s, na.rm = TRUE))

("Rosa")
(min(rosa$diff_s, na.rm = TRUE))
(max(rosa$diff_s, na.rm = TRUE))

```
# Task 3: Deriving Movement Parameters I: Speed

```{r calculate steplength}
boar <- boar |> 
  group_by(TierName) |>
  mutate(steplength = ((E - lead(E))^2 + (N - lead(N))^2)^0.5)
```


```{r calculate speed}
boar <- boar |> 
  group_by(TierName) |>
  mutate(speed_ms = as.numeric(steplength/diff_s))
```

```{r plot speed per step}
ggplot(boar,aes(speed_ms)) +
  geom_histogram( )

ggplot(boar,aes(speed_ms, y = steplength), colour = TierID) +
  geom_point()
```
The histogramm shows us that most steps have very low to no speed. 
The scatter plot might has a beam like structure. each beam might represent the slope as a relation of steplength and speed_ms to the sampling rate. 
